---
output: github_document
editor_options: 
  chunk_output_type: console
bibliography: references.bib
link-citations: yes
linkcolor: Blue
urlcolor: Blue
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
knitr::opts_chunk$set(fig.width=10, fig.height=4)
options(tidyverse.quiet = TRUE)
```


<br>
**<span style="color: red;">Disclaimer</span>**

This package is a work in progress. It has been released to get feedback from 
users that we can incorporate in future releases.

# serosurvey

<!-- badges: start -->
[![DOI](https://zenodo.org/badge/296497540.svg)](https://zenodo.org/badge/latestdoi/296497540)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/serosurvey)](https://cran.r-project.org/package=serosurvey)
<!-- badges: end -->

The goal of `serosurvey` is to gather __Serological Survey Analysis__ functions and workflow templates for __Prevalence Estimation Under Misclassification__.

## Installation

<!-- You can install the released version of serosurvey from [CRAN](https://CRAN.R-project.org) with: -->

``` r
if(!require("remotes")) install.packages("remotes")
remotes::install_github("avallecam/serosurvey")
```

## Example

Three basic examples which shows you how to solve common problems:

```{r example}
library(serosurvey)
```

```{r,echo=FALSE}
# additional
library(tidyverse)
library(srvyr)
library(survey)
library(tictoc)
library(furrr)
library(purrr)
# theme
theme_set(theme_bw())
```

```{r,echo=FALSE}
data(api)

datasurvey <- apiclus2 %>% 
  mutate(survey_all="survey_all") %>% 
  # create variables
  mutate(outcome_one = awards,
         outcome_two = cut(pct.resp,breaks = 2),
         covariate_01 = stype,
         covariate_02 = both)
```

```{r,echo=FALSE}
# tratamiento de stratos con un solo conglomerado
options(survey.lonely.psu = "certainty")

# uu_clean_data %>% count(CONGLOMERADO,VIVIENDA)

# diseño muestral de la encuesta ---------------------------------

design <- datasurvey %>% 
  
  filter(!is.na(outcome_one)) %>% #CRITICAL! ON OUTCOME
  filter(!is.na(pw)) %>% #NO DEBEN DE HABER CONGLOMERADOS SIN WEIGHT
  
  as_survey_design(
    id=c(dnum, snum), #~dnum+snum, # primary secondary sampling unit
    # strata = strata, #clusters need to be nested in the strata
    weights = pw # factores de expancion
  )
```

```{r,echo=FALSE}
# denominadores
covariate_set01 <- datasurvey %>% 
  select(covariate_01,
         #sch.wide,
         #comp.imp,
         covariate_02) %>% 
  colnames()

# numerators within outcome
covariate_set02 <- datasurvey %>% 
  select(#stype,
         #sch.wide,
         #comp.imp,
         covariate_02) %>% 
  colnames()
```

### 1. `survey`: Estimate single prevalences

- From a [`srvyr`](http://gdfe.co/srvyr/) __survey design object__, __`serosvy_proportion`__ estimates:

  + weighted prevalence (`prop`), 
  + total population (`total`),
  + raw proportion (`raw_prop`), 
  + coefficient of variability (`cv`), 
  + design effect (`deff`)

```{r}
serosvy_proportion(design = design,
                   denominator = covariate_01,
                   numerator = outcome_one)
```

```{r,eval=FALSE}
example("serosvy_proportion")
```

### 2. `survey`: Estimate multiple prevalences

- In 
the [Article tab](https://avallecam.github.io/serosurvey/articles/howto-reprex.html) 
we provide a workflow to __estimate multiple prevalences__: 

  + using different set of covariates and outcomes as numerators or denominators,
  + in one single pipe operation

```{r}
# crear matriz
  #
  # set 01 of denominator-numerator
  #
expand_grid(
  design=list(design),
  denominator=c("covariate_01","covariate_02"), # covariates
  numerator=c("outcome_one","outcome_two") # outcomes
  ) %>% 
  #
  # set 02 of denominator-numerator (e.g. within main outcome)
  #
  union_all(
    expand_grid(
      design=list(design),
      denominator=c("outcome_one","outcome_two"), # outcomes
      numerator=c("covariate_02") # covariates
    )
  ) %>% 
  #
  # create symbols (to be readed as arguments)
  #
  mutate(
    denominator=map(denominator,dplyr::sym),
    numerator=map(numerator,dplyr::sym)
  ) %>% 
  #
  # estimate prevalence
  #
  mutate(output=pmap(.l = select(.,design,denominator,numerator),
                     .f = serosvy_proportion)) %>% 
  #
  # show the outcome
  #
  select(-design,-denominator,-numerator) %>% 
  unnest(cols = c(output)) %>% 
  print(n=Inf)
```

#### `learnr` tutorial

- Learn to build this with in a tutorial in Spanish:

```r
# update package
if(!require("remotes")) install.packages("remotes")
remotes::install_github("avallecam/serosurvey")
# install learner and run tutorial
if(!require("learnr")) install.packages("learnr")
learnr::run_tutorial(name = "taller",package = "serosurvey")
```


### 3. `serology`: Estimate prevalence Under misclassification

- We gather __one frequentist approach__ [@ROGAN1978], 
available in different Github repos, that deal with 
misclassification due to an imperfect diagnostic 
test [@Azman2020; @Takahashi2020].
Check the [Reference tab](https://avallecam.github.io/serosurvey/reference/index.html).

- We provide __tidy outputs for bayesian approaches__ developed 
in @Larremore2020unk [here](https://github.com/LarremoreLab/bayesian-joint-prev-se-sp/blob/master/singleSERO_uncertainTEST.R) 
and @Larremore2020kno [here](https://github.com/LarremoreLab/covid_serological_sampling/blob/master/codebase/seroprevalence.R):

- You can use them with [`purrr`](https://purrr.tidyverse.org/) and [`furrr`](https://davisvaughan.github.io/furrr/) to efficiently iterate 
and parallelize this step for __multiple prevalences__.
Check the workflow 
in [Article tab](https://avallecam.github.io/serosurvey/articles/howto-reprex.html).


#### __Known test performance - Bayesian method__

```{r,eval=FALSE}
serosvy_known_sample_posterior(
  #in population
  positive_number_test = 321,
  total_number_test = 321+1234,
  # known performance
  sensitivity = 0.93,
  specificity = 0.975
)
```

```{r,echo=FALSE}
tidy_result <- serosvy_known_sample_posterior(
  #in population
  positive_number_test = 321,
  total_number_test = 321+1234,
  # known performance
  sensitivity = 0.93,
  specificity = 0.975
)

tidy_result_out <-
  tidy_result %>%
  select(summary) %>%
  unnest(cols = c(summary))

tidy_result %>%
 select(posterior) %>%
 unnest(cols = c(posterior)) %>%
 ggplot(aes(x = r1)) +
 geom_histogram(aes(y=..density..),binwidth = 0.0005) +
 geom_density() +
 geom_vline(aes(xintercept=tidy_result_out %>%
                  pull(numeric.mean)),
            color="red",lwd=1) +
 geom_vline(aes(xintercept=tidy_result_out %>%
                  pull(numeric.p05)),
            color="red") +
 geom_vline(aes(xintercept=tidy_result_out %>%
                  pull(numeric.p95)),
            color="red") +
 scale_x_continuous(breaks = scales::pretty_breaks())
```

```{r,eval=FALSE}
example("serosvy_known_sample_posterior")
```

#### __Unknown test performance - Bayesian method__

- The test performance is called _"unknown"_ or _"uncertain"_ when test 
sensitivity and specificity are not known with 
certainty [@Kritsotakis2020; @Diggle2011; @Gelman2020] and 
lab validation data is available with a limited set of samples, 
tipically during a novel pathogen outbreak.

```{r,eval=FALSE,echo=FALSE}
# result_unk <- sample_posterior_r_mcmc_testun(
#   samps = 10000,
#   #in population
#   pos = 692, #positive
#   n = 3212, #total
#   # in lab (local validation study)
#   tp = 670,tn = 640,fp = 202,fn = 74)
```

```{r,eval=FALSE}
serosvy_unknown_sample_posterior_ii(
  #in population
  positive_number_test = 321,
  total_number_test = 321+1234,
  # in lab (local validation study)
  true_positive = 670,
  true_negative = 640,
  false_positive = 202,
  false_negative = 74)
```

```{r,echo=FALSE}
result_unk <- serosvy_unknown_sample_posterior_ii(
  #in population
  positive_number_test = 321,
  total_number_test = 321+1234,
  # in lab (local validation study)
  true_positive = 670,
  true_negative = 640,
  false_positive = 202,
  false_negative = 74)

result_unk %>%
  select(posterior) %>% 
  unnest(posterior) %>%
  rownames_to_column() %>%
  pivot_longer(cols = -rowname,
               names_to = "estimates",
               values_to = "values") %>%
  ggplot(aes(x = values)) +
  geom_histogram(aes(y=..density..),binwidth = 0.0005) +
  geom_density() +
  facet_grid(~estimates,scales = "free_x")
```

```{r,eval=FALSE}
example("serosvy_unknown_sample_posterior")
```

## Contributing

Feel free to fill an issue or contribute with your functions or workflows in a pull request. 

Here are a list of publications with interesting approaches using R:

- @Silveira2020 and @Hallal2020 analysed a serological survey accounting for sampling design and test validity using parametric bootstraping, following @Lewis2012.

- @Flor2020 implemented a lot of frequentist and bayesian methods for test with known sensitivity and specificity. Code is available [here](https://github.com/BfRstats/bayespem-validation-code).

- @Gelman2020 also applied Bayesian inference with hierarchical regression 
and post-stratification to account for test uncertainty
with unknown specificity and sensitivity. 
Here a [case-study](https://github.com/bob-carpenter/diagnostic-testing/blob/master/src/case-study/seroprevalence-meta-analysis.Rmd).

## How to cite this R package

```{r}
citation("serosurvey")
```

## Contact

Andree Valle Campos | 
[`@avallecam`](https://twitter.com/avallecam) |
avallecam@gmail.com

Project Link: https://github.com/avallecam/serosurvey

## Acknowledgements

Many thanks to the Centro Nacional de Epidemiología, Prevención y Control 
de Enfermedades [(CDC Perú)](https://www.dge.gob.pe/portalnuevo/)
for the opportunity to work on this project.

## References

Azman, Andrew S, Stephen Lauer, M. Taufiqur Rahman Bhuiyan, Francisco J Luquero, Daniel T Leung, Sonia Hegde, Jason B Harris, et al. 2020. “Vibrio Cholerae O1 Transmission in Bangladesh: Insights from a Nationally- Representative Serosurvey,” March. https://doi.org/10.1101/2020.03.13.20035352.

Diggle, Peter J. 2011. “Estimating Prevalence Using an Imperfect Test.” Epidemiology Research International 2011: 1–5. https://doi.org/10.1155/2011/608719.

Flor, Matthias, Michael Weiß, Thomas Selhorst, Christine Müller-Graf, and Matthias Greiner. 2020. “Comparison of Bayesian and Frequentist Methods for Prevalence Estimation Under Misclassification.” BMC Public Health 20 (1). https://doi.org/10.1186/s12889-020-09177-4.

Gelman, Andrew, and Bob Carpenter. 2020. “Bayesian Analysis of Tests with Unknown Specificity and Sensitivity.” Journal of the Royal Statistical Society: Series C (Applied Statistics), August. https://doi.org/10.1111/rssc.12435.

Hallal, Pedro C, Fernando P Hartwig, Bernardo L Horta, Mariângela F Silveira, Claudio J Struchiner, Luı́s P Vidaletti, Nelson A Neumann, et al. 2020. “SARS-CoV-2 Antibody Prevalence in Brazil: Results from Two Successive Nationwide Serological Household Surveys.” The Lancet Global Health, September. https://doi.org/10.1016/s2214-109x(20)30387-9.

Kritsotakis, Evangelos I. 2020. “On the Importance of Population-Based Serological Surveys of SARS-CoV-2 Without Overlooking Their Inherent Uncertainties.” Public Health in Practice 1 (November): 100013. https://doi.org/10.1016/j.puhip.2020.100013.

Larremore, Daniel B., Bailey K Fosdick, Kate M Bubar, Sam Zhang, Stephen M Kissler, C. Jessica E. Metcalf, Caroline Buckee, and Yonatan Grad.2020.“Estimating SARS-CoV-2 Seroprevalence and Epidemiological Parameters with Uncertainty from Serological Surveys.” medRxiv, April. https://doi.org/10.1101/2020.04.15.20067066.

Larremore, Daniel B., Bailey K. Fosdick, Sam Zhang, and Yonatan Grad.2020.“Jointly Modeling Prevalence, Sensitivity and Specificity for Optimal Sample Allocation.” bioRxiv, May. https://doi.org/10.1101/2020.05.23.112649.

Lewis, Fraser I, and Paul R Torgerson. 2012. “A Tutorial in Estimating the Prevalence of Disease in Humans and Animals in the Absence of a Gold Standard Diagnostic.” Emerging Themes in Epidemiology 9 (1). https://doi.org/10.1186/1742-7622-9-9.

Rogan, Walter J., and Beth Gladen. 1978. “Estimating Prevalence from the Results of A Screening Test.” American Journal of Epidemiology 107 (1): 71–76. https://doi.org/10.1093/oxfordjournals.aje.a112510.

Silveira, Mariângela F., Aluı́sio J. D. Barros, Bernardo L. Horta, Lúcia C. Pellanda, Gabriel D. Victora, Odir A. Dellagostin, Claudio J. Struchiner, et al. 2020. “Population-Based Surveys of Antibodies Against SARS-CoV-2 in Southern Brazil.” Nature Medicine 26 (8): 1196–9. https://doi.org/10.1038/s41591-020-0992-3.

Takahashi, Saki, Bryan Greenhouse, and Isabel Rodríguez-Barraquer. 2020. “Are SARS-CoV-2 seroprevalence estimates biased?” The Journal of Infectious Diseases, August. https://doi.org/10.1093/infdis/jiaa523.
