---
output: github_document
editor_options: 
  chunk_output_type: console
bibliography: references.bib
link-citations: yes
linkcolor: Blue
urlcolor: Blue
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
knitr::opts_chunk$set(fig.width=10, fig.height=4)
options(tidyverse.quiet = TRUE)
```


<br>
**<span style="color: red;">Disclaimer</span>**

This package is a work in progress. It has been released to get feedback from 
users that we can incorporate in future releases.

# serosurvey

<!-- badges: start -->
[![DOI](https://zenodo.org/badge/296497540.svg)](https://zenodo.org/badge/latestdoi/296497540)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/serosurvey)](https://cran.r-project.org/package=serosurvey)
[![Codecov test coverage](https://codecov.io/gh/avallecam/serosurvey/branch/master/graph/badge.svg)](https://codecov.io/gh/avallecam/serosurvey?branch=master)
[![R-CMD-check](https://github.com/avallecam/serosurvey/workflows/R-CMD-check/badge.svg)](https://github.com/avallecam/serosurvey/actions)
<!-- badges: end -->

The goal of `serosurvey` is to gather __Serological Survey Analysis__ functions and workflow templates for __Prevalence Estimation Under Misclassification__.

## Installation

<!-- You can install the released version of serosurvey from [CRAN](https://CRAN.R-project.org) with: -->

You can install the developmental version of `serosurvey` from 
[GitHub](https://github.com/avallecam/serosurvey) with:

``` r
if(!require("remotes")) install.packages("remotes")
remotes::install_github("avallecam/serosurvey")
```

## Brief description

The current workflow is divided in two steps:

1. `survey`: Estimate multiple prevalences, and
2. `serology`: Estimate prevalence Under misclassification for a device 
with Known or Unknown test performance

## More

- In the 
[Introductory article](https://avallecam.github.io/serosurvey/articles/intro.html)
we provide detailed definitions and references of the methods available.
- In 
the 
[Workflow article](https://avallecam.github.io/serosurvey/articles/howto-reprex.html) 
we provide a reproducible example with this package. 


## Contributing

Feel free to fill an issue or contribute with your functions or workflows in a pull request. 

## Code of Conduct

Please note that the epihelper project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.


## Contact

Andree Valle Campos | 
[`@avallecam`](https://twitter.com/avallecam) |
avallecam@gmail.com

Project Link: https://github.com/avallecam/serosurvey

## Acknowledgements

Many thanks to the Centro Nacional de Epidemiología, Prevención y Control 
de Enfermedades [(CDC Perú)](https://www.dge.gob.pe/portalnuevo/)
for the opportunity to work on this project.

## How to cite this R package

```{r}
citation("serosurvey")
```
