---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
knitr::opts_chunk$set(fig.width=10, fig.height=4)
options(tidyverse.quiet = TRUE)
```
# serosurvey

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/serosurvey)](https://cran.r-project.org/package=serosurvey)
<!-- badges: end -->

The goal of serosurvey is to provide a centralized R package of Serological Survey Analysis For Prevalence Estimation Under Misclassification

## Installation

<!-- You can install the released version of serosurvey from [CRAN](https://CRAN.R-project.org) with: -->

``` r
if(!require("devtools")) install.packages("devtools")
devtools::install_github("avallecam/serosurvey")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(serosurvey)
## basic example code
```

### Estimate prevalence from a survey design

```{r,eval=FALSE}
example("serosvy_proportion")
```

### Consider misclassification due to an imperfect diagnostic test

#### Known test - Bayesian method

```{r,eval=FALSE}
example("serosvy_known_sample_posterior")
```

#### Unknown test - Bayesian method

```{r,eval=FALSE}
example("serosvy_unknown_sample_posterior")
```

## Workflow

> pending

## Contribute

Feel free to contribute with your functions or workflows. 
Here are two articles with interesting approaches using R.

- Silveira et al (2020) in the strategy 3 analysed the survey accounting for sampling design and test validity using parametric bootstraping:

> Silveira, M. F., Barros, A. J., Horta, B. L., Pellanda, L. C., Victora, G. D., Dellagostin, O. A., ... & Mesa, J. M. (2020). Population-based surveys of antibodies against SARS-CoV-2 in Southern Brazil. Nature Medicine, 26(8), 1196-1199. doi: https://doi.org/10.1038/s41591-020-0992-3

- Flor et al. (2020) implemented a lot of frequentist and bayesian methods for test with known sensitivity and specificity 

> Flor, M., Weiβ, M., Selhorst, T., Müller-Graf, C., & Greiner, M. (2020). Comparison of Bayesian and frequentist methods for prevalence estimation under misclassification. doi: https://doi.org/10.1186/s12889-020-09177-4

## Citation

```{r}
citation("serosurvey")
```

